### 2019-07-16 manuscript/discussion.tex
Rosana edited the whole thing

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
%
Species are composed of vast assemblages of variable traits, many traits are heritable, and many traits could affect the propensity of species to perish or multiply \citep{lewontin_1970}.
Examining the effects of complex trait combinations on lineage diversification, however, remains challenging.
%
Focusing first on ploidy and then on breeding system, we found that considering each trait in isolation provides an incomplete story.
Considering them together, and in conjunction with another hidden factor, provides a fuller picture of macroevolutionary dynamics within Solanaceae.
We hope our work serves as an example of how phylogenetic comparative methods can be used to disentangle the contributions of interacting traits to heterogeneous lineage diversification.

@@

% E: should this results summary be much shorter?
Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}. 
We obtain a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification is not clear when we also allow that a hidden factor could affect diversification as well (\cref{figure:netdivall}B).
%
Previous analyses of breeding system in this family showed that SI species are associated with higher diversification rates than their SC counterparts (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
Our new analyses with a hidden trait show that this pattern persists on different backgrounds.
One on background, however, SC species experience positive net diversification (\ie not a dead end; cf.\ \citealt{igic_2013}) that exceeds the diversification rate of SI species on the other background (\cref{figure:netdivall}D).
%
Furthermore, our models of the joint macroevolution of ploidy and breeding system provided better statistical fits and revealed how the trait combinations associate with diversification.
We find that the higher overall high net diversification rate of diploids is dominated by SI diploids, while SC diploids have a lower diversification rate that is comparable to polyploids, which are also SC (\cref{figure:netdivall}E).
Thus, breeding system seems to provide the dominant explanation of diversification differences, with ploidy providing a secondary effect within SC species.

@@

Throughout our numerous model comparisons, models that included a hidden trait emerged as significantly better than those that did not. % supptable ref
This is entirely consistent with the expectation that many processes can affect speciation and extinction, far beyond the one or two traits of particular interest to us.
A natural follow-up question is, what is this hidden trait?
Our results show that to some extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa.
But there is also strong statistical support for a hidden trait beyond ploidy and breeding system.
One temptation is to propose specific possible traits to add into the analysis.
For example, the rapidly-diversifying clade of mostly SC species within \emph{Solanum} suggests that geography may play a role. % Boris, please check!  Australian?  fixme
%     but don't be circular, test in different clade
% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
Alternatively, we could view the hidden trait as merely a statistical trick, providing an easy way to model extra heterogeneity without directly representing a real trait.

@@

Polyploidization is known to be common in plants, but the commonness of the reverse process---diploidization---is under active investigation.
%
In our analyses, ignoring diploidization if it is common could cause us to underestimate the polyploidization rate, overestimate of the diploid diversification rate, or otherwise confound our conclusions.
We therefore included a slate of models with a diploidization parameter, and they showed that our main conclusions are robust to this process.
These models also suggested modest statistical support for diploidization occuring within Solanaceae \cref{supptableFIXME}, although our estimates of the diploidization rate were highly uncertain.
Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus chromosome multiplicity that we primarily relied on) would be needed for more reliable conclusions from an analysis such as ours.
%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization might be a potential issue due to ploidy misclassification.  
%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with comparative phylogenetic models. 
%B: yielding to %R here. I still think there is overwhelming support for this statement, and few of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization, rendering "relatively rare" obsolete or invalid. (P.S. Rare not the same as unimportant!)
%B: was: Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.

>>>>

%B:  at least two surprising outcomes: (1)  D and P on background of C, and (2) SC not as dead of an end as it was in goldberg2010, if P is partitioned out (can be framed as kind of path-dependence, i.e. qIC path different than rho path, or same as (1), so that I and C are less different on the background of D.
%
Species are composed of vast assemblages of variable traits, many traits are heritable, and many traits could affect the propensity of species to perish or multiply \citep{lewontin_1970}.
Examining the effects of complex trait combinations on lineage diversification, however, remains challenging.
%
Focusing first on ploidy and then on breeding system, we found that considering each trait in isolation provides an incomplete story.
Considering them together, and in conjunction with another hidden factor, provides a fuller picture of macroevolutionary dynamics within Solanaceae.
We hope our work serves as an example of how phylogenetic comparative methods can be used to disentangle the contributions of interacting traits to heterogeneous lineage diversification, and how to statistically argue for increasing complexity in diversification modeling.

@@

% E: should this results summary be much shorter?
Previous analyses of the effects of ploidy on diversification showed that diploids had greater net diversification rates than polyploids across many angiosperm clades \citep{mayrose_2011, mayrose_2015}. 
We obtained a similar outcome when examining ploidy alone in Solanaceae (\cref{figure:netdivall}A), but a consistent effect of ploidy on diversification was not clear when we added a  hidden factor linked to diversification(\cref{figure:netdivall}B).
%
Previous analyses of breeding system in this family showed that SI species were associated with higher diversification rates than their SC counterparts (\citealt{goldberg_2010}; \cref{figure:netdivall}C).
Our new analyses with a hidden trait showed that this pattern persists even under the presences of an unobserved trait linked to diversification.
Under one of the hidden states, however, SC species experienced positive net diversification (\ie not a dead end; cf.\ \citealt{igic_2013}) that exceeded the diversification rate of SI species under the second hidden state (\cref{figure:netdivall}D).
%
Furthermore, our models of the joint macroevolution of ploidy and breeding system that were statistically preferred revealed how the trait combinations associate with diversification.
We found that the highest overall net diversification rate of diploids was dominated by SI diploids, while SC diploids had a lower diversification rate that overlapped with the net diversification of polyploids, which are also SC (\cref{figure:netdivall}E).
Thus, breeding system seems to provide the dominant explanation of diversification differences, with ploidy providing a secondary effect within SC species.

@@

Throughout our numerous model comparisons, models that included a hidden trait emerged as significantly better than those that did not (\cref{supptable:testaddhidden}). 
This is entirely consistent with the expectation that many processes can affect speciation and extinction, far beyond the one or two traits of particular interest to us.
A natural follow-up question is, what is this hidden trait?
Our results show that to some extent, breeding system functions as a hidden trait in the ploidy-centered analysis, and vice versa (\cref{suppfigure:lumpedDP,suppfigure:lumpedIC}).
But there is also strong statistical support for a hidden trait beyond ploidy and breeding system (\cref{supptable:testaddhidden}).
One temptation is to propose specific possible traits to add into the analysis.
For example, the rapidly-diversifying clade of mostly SC species within \emph{Solanum} suggests that geography may play a role. % Boris, please check!  Australian?  fixme
%     but don't be circular, test in different clade
% alpha is always low: new supp fig showing how A & B are inferred on the tree?  at least for ID/CD/CP+A/B model
%Here Rosana adds the value of adding traits with the lumped models
When analyzing the value of adding more traits to a complex model, our suggestion is to opt for the use of lumped models (\cref{figure:lumped,table:lumped}), and verify that adding a trait is statistically preferred.
In the absence of additional information, the hidden states can be view as merely a statistical trick, providing an easy way to model extra heterogeneity without directly representing a specific trait.
In our analyses, the transition rates between and within hidden traits did not affect the direction of the net diversification difference. However, we found that it is important to allow for the hidden state transition rates to be asymmetrical since those models have a better fit over models that fix the hidden state transition rates to be equal (\cref{suppfigure:asymmetric,supptable:asymmetry}).

@@

Polyploidization is known to be common in plants, but the commonness of the reverse process---diploidization---is under active investigation \citep{dodsworth_2015,mandakova_2018}.
%
Ignoring diploidization if it is common, could cause us  for models that ignore it to underestimate the polyploidization rate and overestimate of the diploid diversification rate leading to incorrect conclusions.
Therefore, we included a slate of models with a diploidization parameter, and they showed that our main conclusions are robust to this process (\cref{suppfigure:alldip}).
These models also suggested modest statistical support for diploidization occuring within Solanaceae (\cref{supptable:testdiploidization}), although our estimates of the diploidization rate were highly uncertain.
Furthermore, additional lines of evidence for classifying species as diploid or polyploid (beyond the genus chromosome multiplicity that we primarily relied on) would be needed for more reliable conclusions from an analysis such as ours.
%In the cases where ploidy level was assigned by us based chromosome multiplicity at the genus level, estimating diploidization might be a potential issue due to ploidy misclassification.  
%R: By adding diploidization to models of polyploidy linked to diversification it is possible to recover this complicated scenario and to reconcile genomic evidence with comparative phylogenetic models. 
%B: yielding to %R here. I still think there is overwhelming support for this statement, and few of the cited papers show convincingly that, for example, diploidization happens within five orders of magnitude compared with polyploidization, rendering "relatively rare" obsolete or invalid. (P.S. Rare not the same as unimportant!)
%B: was: Diploidization is widely considered to be relatively rare, compared with polyploidization \citep{husband_2013}, but flowering plant lineages are thought to have experienced at least one round of polyploidization in their evolutionary history \citep{soltis_2015}.

### 2019-07-16 manuscript/introduction.tex
Rosana edited the whole thing

Species accumulate across the tree of life at different rates. 
One possible explanation for this phenomenon is that the various traits differentially affect rates of diversification. %B: note: traits and character states are synonymous. I've edited the rest of the manuscript so that "trait" is retained where ambiguity does not matter, but where directly contrasted (vs.: as here), I changed the wording.
Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification traits. 
%B: was: "phenotypic data and methodological advances"
Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
One difficulty is that speciation and extinction likely do not depend on a single character, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
Consequently, examining the association of only one character with diversification patterns can be misleading. 
Here, we embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors.
% E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
% Could also state that we newly test whether adding one more trait is worthwhile.
%B: I think you are correct, but it introduces a new concern: we now say N=1 is too few, so not why try N>2 characters? Why is N=2 the Goldilocks number? I didn't change anything but it's clear fodder for Discussion.

Polyploidization is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
Increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
%B: The contrast below was between: "Polyploids have been documented..." vs. "Other studies find...". I modified it to: "Polyploids were generally found..." vs. "more recent studies ..."; If you feel that "Many studies" or "Recent studies" is better, please alter as appropriate.
%B: Separately, "common" and "numerous" are synonymous as used here (I removed "common" because that term also has the "unrefined" meaning).
Polyploids were generally found to have an overall lower diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
Many recent genomic studies, however, find numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants. 
Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} to include transitions between diploid and polyploid states. 
We also allow for heterogeneous diversification rates within each ploidy level, which could examine whether, for example, diploids with ancestral polyploidy diversify more rapidly than diploids without it.

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are also remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{barrett2013}. %B: note: we use "remarkably" to describe the frequency of both characters; hence "also."
In particular, a variety of genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and loss of such mechanisms, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2019}.
Although these findings suggest a consistent macroevolutionary role of breeding system, it is unlikely to be the sole character determining lineage diversification.
We thus investigate the relationship of breeding system to speciation and extinction rates in the context of ploidy and other background factors.

@@

%B: redundant with "very powerful" below rm: Examining these questions requires powerful phylogenetic comparative methods.
In the present study, we employ an extended framework of state speciation and extincion models (SSE), which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.
%Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}. We follow by 
fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor that is changing the diversification pattern \citep{beaulieu_2016}.
We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually linked to the diversification process.
Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character. With the joint trait model without hidden state, we also quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
Later, we propose new models that aggregate states in the ploidy and breeding system models to make them equivalent to the single trait models.
The new `lumped' models allow us to compare whether increasing model complexity from one trait to two traits significantly improves our understanding of the diversification process.
%B: I altered this line, but it is still a bit redundant given previous sentence.
%Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
%First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
%Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
%
% E: say something here about testing whether it's worth adding in another real trait?
Then, we extend all the models involving ploidy, to investigate the potential effects of including diploidization.
Finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.

Our results highlight the importance of considering non-additive effects of traits on net diversification rates, especially when there are strong biologically-driven correlations among them.

>>>>

Species accumulate across the tree of life at different rates. 
One possible explanation for this phenomenon is that the various traits differentially affect rates of diversification. %B: note: traits and character states are synonymous. I've edited the rest of the manuscript so that "trait" is retained where ambiguity does not matter, but where directly contrasted (vs.: as here), I changed the wording.
Dramatic increases in phylogenetic and phenotypic data along with methodological advances have greatly accelerated the search for traits that influence diversification. 
%B: was: "phenotypic data and methodological advances"
Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a challenge \citep[\eg][]{maddison_2015, rabosky_2015, beaulieu_2016, rabosky_2017}. 
One difficulty is that speciation and extinction likely do not depend on a single character, so the biological and environmental contexts in which traits occur can lead to complex interactions that affect lineage diversification \citep{beaulieu_2016, caetano_2018, herrera_2018}.
Consequently, examining the association of only one character with diversification patterns can be misleading. 
We embrace this challenge by jointly investigating two characters thought to influence speciation and extinction rates---ploidy level and breeding system---while allowing for their interactions and other confounding factors to change diversification.
% E: We need a "Here, we..." here.  Otherwise, we've just told the reader "not one trait!" and then launch into one trait.
% Could also state that we newly test whether adding one more trait is worthwhile.
%B: I think you are correct, but it introduces a new concern: we now say N=1 is too few, so not why try N>2 characters? Why is N=2 the Goldilocks number? I didn't change anything but it's clear fodder for Discussion.

Polyploidization is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
The widespread prevalence of variation in ploidy has long been considered a salient feature of flowering plant lineages \citep{stebbins1938}. 
Increase in ploidy can alter many traits and affect a variety of evolutionary and ecological processes \citep{ramsey_2002, sessa_2019}.
%B: The contrast below was between: "Polyploids have been documented..." vs. "Other studies find...". I modified it to: "Polyploids were generally found..." vs. "more recent studies ..."; If you feel that "Many studies" or "Recent studies" is better, please alter as appropriate.
%B: Separately, "common" and "numerous" are synonymous as used here (I removed "common" because that term also has the "unrefined" meaning).
Polyploids were  found to have an overall lower net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
However, many recent genomic studies,  have found numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that whole genome duplications may have played an important macroevolutionary role driving innovation and diversification in plants. 
Evidence of paleo-polyploidy within the genomes of diploid extant plants also implies pervasive diploidization, the return of polyploids to the diploid state, throughout the angiosperm phylogeny \citep{soltis_2015, dodsworth_2015}.
Our analyses re-examine the association between ploidy and lineage diversification by extending the approach of \citet{mayrose_2011, mayrose_2015} by including transitions from polyploid to diploid states and potential unobserved factors affecting diversification patterns.
%R- No we don't do the below thing and it was confusing that's what Jacob Landis did in 2018
%We also allow for heterogeneous diversification rates within each ploidy level, which could examine whether, for example, diploids with ancestral polyploidy diversify more rapidly than diploids without it.

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are also remarkably common and crucially affect the distribution and amount of genetic variation in populations \citep{barrett2013}. %B: note: we use "remarkably" to describe the frequency of both characters; hence "also."
In particular, a variety of genetic self-incompatibility (SI) systems cause plants to reject their own pollen, and loss of such mechanisms, yielding self-compatibility (SC), is one of the most replicated transitions in flowering plant evolution \citep{stebbins1974, igic_2008}.
Previous analyses reported higher rates of diversification for SI than for SC lineages in Solanaceae \citep{goldberg_2010}. 
Similarly, heterostylous SI lineages in Primulaceae seem to diversify faster \citep{devos2014}, as do outcrossing lineages in Onagraceae \citep{freyman_2019}.
Although these findings suggest a consistent macroevolutionary role of breeding system, it is unlikely to be the sole character determining lineage diversification.
We investigate the relationship of breeding system to speciation and extinction rates in the context of ploidy and other unobserved factors.

@@

%B: redundant with "very powerful" below rm: Examining these questions requires powerful phylogenetic comparative methods.
In the present study, we employ an extended framework of state speciation and extincion models (SSE), which simultaneously models transitions between the discrete states of a trait and different rates of speciation and extinction associated with each of those states \citep{maddison_2007, fitzjohn_2012}.
All the models proposed for this study are coded as graphical models and Bayesian inference was performed over the proposed models using RevBayes software \citep{hoehna_2016}.
%Although potentially very powerful, this approach can be prone to reporting a false association between a trait and lineage diversification when the data contain additional, unmodeled diversification rate heterogeneity \citep{maddison_2015, rabosky_2015}.
We start by fitting binary state speciation and extinction models to each ploidy and breeding system independently \citep{maddison_2007}. We follow by 
fitting models that incorporate hidden states, which reduce the chance that an effect of one trait on diversification would be mistaken for the effect of an unobserved factor \citep{beaulieu_2016}.
We compare the proposed models against their character independent counterparts \citep{beaulieu_2016} to investigate whether ploidy or breeding system are actually informative of the diversification process.
Next, we model ploidy and breeding system jointly to assess their combined influence on diversification, with or without an additional hidden character. Using thw ploidy and breeding system model  without hidden state, we quantify the relative contributions of the two pathways from SI diploids to SC polyploids.
Later, we propose new models that aggregate states in the ploidy and breeding system models to make them equivalent to the binary state models.
The new `lumped' models allow us to compare whether increasing model complexity from one trait to two traits significantly improves our understanding of the diversification process.
%B: I altered this line, but it is still a bit redundant given previous sentence.
%Our analyses evaluate combinations of macroevolutionary models to parse out the associations of ploidy and breeding system on speciation and extinction rates in Solanaceae, while simultaneously accounting for other processes that influence diversification.
%First, we present models in which the effects of ploidy and breeding system are considered separately and compare the inferences to previously published results. 
%Second, we add a hidden character to each model and investigate if state-dependent diversification in focal character retains its explanatory power, or whether the added hidden character obscures it.
%
% E: say something here about testing whether it's worth adding in another real trait?
Furthermore, we extend all the models involving ploidy, to investigate the potential effects of including diploidization.
And finally, we explore models of diversification using asymmetric rates between and within hidden states to investigate the sensitivity of the statistical inference to model assumptions.

Our results highlight the importance of considering non-additive effects of traits on net diversification rates  under the presence of unobserved factors, in order to detect strong biologically-driven processes dictating the diversification patterns.
